---
title: "Motivation and Application"
author: "lindbrook"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Motivation and Application}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE #,
  # comment = "#>"
)

library(deldirPolygons)
library(cholera)
```

Using just the locations of sites or landmarks, Voronoi tessellation partitions a space into cells or tiles that represent those sites' neighborhoods (i.e., their catchment or service areas). This package adds to the functionality of the 'deldir' package by computing the vertices of those tiles so that we can leverage functions that use polygons. Doing so simplifies tasks like color coding neighborhoods or counting elements of interest within neighborhoods.

As an example, I use data from John Snow's map of the 1854 cholera outbreak in the Soho area London.

## Coloring Tiles

```{r coloring, fig.align = "center", fig.width = 5, fig.height = 5, echo = TRUE, eval = TRUE}
# compute vertices of Voronoi tiles
vertices <- deldirPolygons(sites = cholera::pumps,
  rw.data = cholera::roads)

# define colors, plot map, and color code fatalities
snow.colors <- grDevices::adjustcolor(cholera::snowColors(), alpha.f = 1/3)
cholera::snowMap(add.cases = FALSE)
cholera::addNeighborhoodCases(metric = "euclidean")

# plot color coded polygons
invisible(lapply(seq_along(vertices), function(i) {
  polygon(vertices[[i]], col = snow.colors[[i]])
}))
```

## Counting Observations in Tiles

To count the number of cases within each neighborhood, we can use sp::point.in.polygon().

```{r counting, echo = TRUE, eval = TRUE}
# compute vertices of Voronoi tiles
vertices <- deldirPolygons(sites = cholera::pumps,
  rw.data = cholera::roads)

# locations of the 578 fatalities in Soho
cases <- cholera::fatalities.unstacked

# count fatalities within each polygon (neigborhood)
census <- lapply(vertices, function(tile) {
  sp::point.in.polygon(cases$x, cases$y, tile$x, tile$y)
})

# ID the 13 water pumps
names(census) <- paste0("p", cholera::pumps$id)

# count of fatalities by neighborhood
vapply(census, sum, integer(1L))
```

## Counting Observations in Triangles

To count the number of cases within each triangle:

```{r counting_triangles, echo = TRUE, eval = TRUE}
# compute vertices of Delauny triangles
vertices <- deldirPolygons(sites = cholera::pumps,
  rw.data = cholera::roads, type = "triangles")

# locations of the 578 fatalities in Soho
cases <- cholera::fatalities.unstacked

# count fatalities within each triangle
census <- lapply(vertices, function(tile) {
  sp::point.in.polygon(cases$x, cases$y, tile$x, tile$y)
})

# ID triangles
names(census) <- paste0("t", seq_along(vertices))

# count of fatalities by triangle
vapply(census, sum, integer(1L))
```
